---
title: "STAT433: Homework 1"
author: "Doy Kim"
date: "September 27, 2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)

```

# Prob 1

How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

## Missing dep_time

```{r P1a}
df = flights

(missing_dep_time = df %>%
  filter(is.na(dep_time) == TRUE) %>%
  summarize(nn = n()))
```

> `r missing_dep_time` flights have a missing dep_time.

## Other variables with missing values

```{r}
colSums(is.na(df))
  
```

> Beside `dep_time`, the following variables contain missing values: `dep_dely`, `arr_time`, `arr_delay`, `tailnum`, `air_time`

## Interpretation

```{r}
df.missing = df %>%
  filter(is.na(dep_time) == TRUE,
         is.na(dep_delay) == TRUE,
         is.na(arr_time) == TRUE,
         is.na(arr_delay) == TRUE,
         is.na(air_time) == TRUE)

head(df.missing, 10)


```

> Note that I ignored missing values in `tailnum` since it is an aircraft identifier, not a characteristic of a specific flight. 
 Having simultaneous missing values in `dep_time`, `dep_delay`, `arr_time`, `arr_delay`, and `air_time` is likely to mean that the filghts were scheduled, but ended up not taking place.

# Prob 2

Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
df_new_time = df %>%
  select(dep_time, sched_dep_time, hour) %>%
  mutate(dep_hour = floor(dep_time/100),
         new_dep_time = dep_hour*60 + (dep_time - dep_hour*100),
         sched_hour = hour,
         new_sched_time = sched_hour*60 + (sched_dep_time - sched_hour*100)) %>%
  select(dep_time, new_dep_time, sched_dep_time, new_sched_time)

head(df_new_time, 10)

```

# Prob 3

Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()

## Is there a pattern?

```{R}
# canceled
(
  df.canceled = df %>% 
  group_by(day) %>% 
  summarize(canceled = sum(is.na(dep_time) == TRUE),
            total = n(),
            p = canceled/total,
            m_delay = mean(dep_delay + arr_delay, na.rm = TRUE))
 )

# Graph
df.canceled %>%
  ggplot(aes(x = day, y = p)) +
  geom_bar(stat = "identity")

# Pattern? Random?
df.chi = df.canceled %>%
  mutate(non_canceled = total - canceled) %>%
  select(canceled, non_canceled)

chisq.test(df.chi)
```

> The bar graph indicates that around filght cancellation is the most frequent aroung the 10th of a month. The chi-squared test's result also suggest that the cancellation pattern across different days is not random. However, a more thorough investigation is required to specify the characteristics of the pattern.

## Is the proportion of canceled flights related to the average delay?

```{r}
lm_out = lm(data = df.canceled, m_delay ~ p)
summary(lm_out)

df.canceled %>%
  ggplot(aes(x = p, y = m_delay)) +
  geom_point() +
  geom_smooth(method = "lm")
```

>The regression results and the graph suggest that the average delay in a day (`m_delay`) and the proportion of canceled flights in a day (`p`) are significantly associated.